---
title: "Thesis_HurdelModels"
author: "Amanda Hastings"
date: '2022-03-30'
output: html_document
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
#Load libraries 
library(tidyverse)
library(dplyr)
library(RColorBrewer)
library(emmeans)
library(multcomp)
library(multcompView)
library(ggthemes)
library(gh)
library(pscl)
```


```{r}
#Read in data 
ThesisData<-read.csv("2021.RDataHastings.csv")
```



# Futher formatting for ThesisData  
```{r}
#Add AspectCategory, XMCategory, allseedlings_total columns 
#Format some column names and levels
ThesisData <- ThesisData %>% 
  mutate("AspectCategory" = case_when(aspect >=45 & aspect <= 134 ~ "East",
         aspect >= 135 & aspect <=224 ~ "South", 
         aspect >=225 & aspect <= 314 ~ "West", 
         TRUE ~ "North"))%>% 
  mutate("XMCategory" = case_when(AspectCategory == "East"|AspectCategory == "North" ~ "Mesic", 
                                  AspectCategory == "West"|AspectCategory == "South" ~ "Xeric"))%>% 
  rowwise()%>%
  mutate("allseedlings_total" = sum(conifer_total,deciduous_total)) %>% 
  mutate(forest_type = factor(
    forest_type,
    levels = c("Lower Montane", "Upper Montane", "Subalpine")
  )) %>% 
  rename("slope"=slope....)


ThesisData[,c(1,2,3,4,5,6,7,8,9,10,11,75:ncol(ThesisData))]
```


# Plant Cover Long Data
```{r}
# Convert plant cover data to long format
PlantCoverLong <- ThesisData %>%
  dplyr::select(forest_type, burn_sev_category, gm_cover:tr_cover) %>%
  tidyr::pivot_longer(gm_cover:tr_cover,
               names_to = "FunctionalGroup",
               values_to = "PercentPlantCover") %>%
  mutate(FunctionalGroup = factor(FunctionalGroup)) %>%
  mutate(forest_type = factor(
    forest_type,
    levels = c("Lower Montane", "Upper Montane", "Subalpine")
  ))
```


# Plant Cover Long Data plus Binary Column
```{r}
# Round percent cover to four decimal places 
PlantCoverLong$PercentPlantCover <- round(PlantCoverLong$PercentPlantCover,4)


PlantCoverBinary <- PlantCoverLong %>% 
  mutate("BinaryCover" = case_when(PercentPlantCover == 0.0000 ~ 0,
         PercentPlantCover != 0.0000 ~ 1))
```


# Summary Statistics All Plant Cover 
```{r}
# Create summary statistics table for plant cover including number of zeroes
StatsPlantCover <- PlantCoverBinary %>% 
  group_by(burn_sev_category, forest_type, FunctionalGroup)%>% 
  summarize(n=n(), 
            mean=mean(PercentPlantCover), 
            sd=sd(PercentPlantCover), 
            n_zero=sum(BinaryCover==0)) 
StatsPlantCover
```


# Plant Cover Nonzeroes 
```{r}
PlantCoverNonZero <- PlantCoverBinary %>% 
  filter(BinaryCover == 1)
```


# Summary Statistics Nonzero Plant Cover 
```{r}
StatsPlantNonZero <- PlantCoverNonZero %>% 
  group_by(burn_sev_category, forest_type, FunctionalGroup)%>% 
  summarize(n=n(), 
            mean=mean(PercentPlantCover), 
            sd=sd(PercentPlantCover))
```


#-----------------------Understory cover analyses-------------------------------


# Shrub Cover Logistic Regression 
```{r}
PlantCoverBinary$BinaryCover <- as.factor(PlantCoverBinary$BinaryCover)

ShrubCoverBinary <- PlantCoverBinary %>% 
  filter(FunctionalGroup == 'sh_cover')

LogisFitSH <- glm(BinaryCover ~ burn_sev_category * forest_type, data=ShrubCoverBinary, family= "binomial" )

library(car)
library(emmeans)
Anova(LogisFitSH,type=3)
emmeans(LogisFitSH,pairwise~forest_type,type="response")
```


# Interaction Plot Shrub Cover
```{r}
ggplot(ShrubCoverBinary,aes(x=forest_type,y=PercentPlantCover,group=burn_sev_category, color=burn_sev_category))+ 
      geom_smooth()
```


# Nonzero Shrub Cover Wide Data
```{r}
ShrubNonZeroWide <- ThesisData %>%
  dplyr::select(ï..PlotID,forest_type, burn_sev_category, gm_cover:tr_cover)%>% 
  filter(sh_cover != 0.000000)
```


# Nonzero Shrub Cover ANOVA - Log Transformed 
```{r}
# Nonzero Shrub ANOVA
hist(ShrubNonZeroWide$sh_cover)
LMFitSH_NZ <-
  lm(log(sh_cover) ~ burn_sev_category * forest_type, data = ShrubNonZeroWide)
anova(LMFitSH_NZ)
par(mfrow = c(1, 2))
plot(LMFitSH_NZ, which = 1:2)
```


# Tree Cover Logistic Regression 
```{r}
TreeCoverBinary <- PlantCoverBinary %>% 
  filter(FunctionalGroup == 'tr_cover')

LogisFitTR <- glm(BinaryCover ~ burn_sev_category * forest_type, data=TreeCoverBinary, family= "binomial" )

Anova(LogisFitTR,type=3)
emmeans(LogisFitTR,pairwise~forest_type,adjust="none",type="response")
```


# Nonzero Tree Cover Wide Data
```{r}
TreeNonZeroWide <- ThesisData %>%
  dplyr::select(ï..PlotID,forest_type, burn_sev_category, gm_cover:tr_cover)%>% 
  filter(tr_cover != 0.000000)
```


# Nonzero Tree Cover ANOVA - Log Transformed
```{r}
# Nonzero Tree ANOVA
hist(TreeNonZeroWide$tr_cover)
LMFitTR_NZ <-
  lm(log(tr_cover) ~ burn_sev_category * forest_type, data = TreeNonZeroWide)
anova(LMFitTR_NZ)
par(mfrow = c(1, 2))
plot(LMFitTR_NZ, which = 1:2)
```


#------------Conifer/decid analyses---------------------------------------------


# Seedling Total Densities Data
```{r}
# Covert seedlings data to long format
SeedlingDensTotal <- ThesisData %>%
  dplyr::select(ï..PlotID,forest_type, burn_sev_category, deciduous_total, conifer_total)

```


# Nonzero Seedling Densities Data
```{r}
SeedlingDensBinary <- SeedlingDensTotal %>% 
  mutate("BinaryDecid" = case_when(deciduous_total == 0.0000 ~ 0,
         deciduous_total != 0.0000 ~ 1)) %>% 
  mutate("BinaryConifer" = case_when(conifer_total == 0.0000 ~ 0,
         conifer_total != 0.0000 ~ 1))
```


## Deciduous Density

# Deciduous Logistic Regression
```{r}
LogisFitDecid <- glm(BinaryDecid ~ burn_sev_category * forest_type, data=SeedlingDensBinary, family= "binomial" )

Anova(LogisFitDecid)
```


# Deciduous Nonzero Densities Data 
```{r}
NonZeroDecid <- SeedlingDensTotal %>% 
  filter(deciduous_total != 0.000000)
  
```


```{r}
# anova(LMFitTR_NZ)
# par(mfrow = c(1, 2))
# plot(LMFitTR_NZ, which = 1:2)
```


## Conifer Density 

# Conifer Logistic Regression
```{r}
LogisFitConifer <- glm(BinaryConifer ~ burn_sev_category * forest_type, data=SeedlingDensBinary, family= "binomial" )

Anova(LogisFitConifer)
```


# Conifer Nonzero Densities Data 
```{r}
# NonZeroConifer <- SeedlingDensTotal %>%
```





