---
title: "ThesisProject_2021"
author: "Amanda Hastings"
date: "3/10/2022"
output: html_document
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
#Load libraries 
library(tidyverse)
library(dplyr)
library(RColorBrewer)
library(emmeans)
library(multcomp)
library(multcompView)
library(ggthemes)
library(gh)
library(pscl)
```


```{r}
#Read in data 
ThesisData<-read.csv("2021.RDataHastings.csv")
```


# Plant Cover Long Data
```{r}
# Convert plant cover data to long format
PlantCoverLong <- ThesisData %>%
  dplyr::select(forest_type, burn_sev_category, gm_cover:tr_cover) %>%
  tidyr::pivot_longer(gm_cover:tr_cover,
               names_to = "FunctionalGroup",
               values_to = "PercentPlantCover") %>%
  mutate(FunctionalGroup = factor(FunctionalGroup)) %>%
  mutate(forest_type = factor(
    forest_type,
    levels = c("Lower Montane", "Upper Montane", "Subalpine")
  ))
```


# Plant Cover Plot
```{r}
# Facet wrap by forest type - plant cover plot
PlantCoverPlot <- ggplot(
  PlantCoverLong,
  aes(x = FunctionalGroup,
      y = PercentPlantCover,
      fill = burn_sev_category)
) +
  geom_boxplot() + ylim(0, 0.6) +
  facet_wrap(~ forest_type) +
  theme_classic() +
  labs(x = "Functional Group",
       y = "Percent Cover",
       fill = "Burn Severity",
       title = "Understory Plant Cover") +
  theme(legend.position = "bottom",
        legend.key.width = unit(1, "cm")) +
  theme_few()+
  scale_fill_manual(
    name = "Burn Severity",
    labels = c("Low", "High"), values=c("#fd8d3c","#bd0026"))+
  theme(axis.text.x = element_text(
    color = "black",
    size = 9,
    angle = 30,
    vjust = .8,
    hjust = .8
  )) +
  scale_x_discrete(labels = c("Graminoids", "Forbs", "Shrubs", "Seedlings"))
PlantCoverPlot 
```


#-----------------------Understory cover analyses-------------------------------


# Three-way ANOVA for Plant Cover by Functional Group
```{r}
#Sqrt transform for percent plant cover
LMFitAllPlants <- lm(sqrt(PercentPlantCover) ~ forest_type*factor(FunctionalGroup)*burn_sev_category, data= PlantCoverLong)
anova(LMFitAllPlants)
par(mfrow = c(1, 2))
plot(LMFitAllPlants, which = 1:2)
```


# Tukey pairs for Three-way ANOVA Plant Cover
```{r}
emoutAllPlants <- emmeans(LMFitAllPlants, ~ forest_type*factor(FunctionalGroup)*burn_sev_category)
emoutAllPlants
pairs(emoutAllPlants)
cld(emoutAllPlants)
```


# Graminoid ANOVA
```{r,echo=FALSE}
# # Graminoid ANOVA
# hist(ThesisData$gm_cover)
# LMFitGM <-
#   lm(gm_cover ~ burn_sev_category * factor(forest_type), data = ThesisData)
# anova(LMFitGM)
# par(mfrow = c(1, 2))
# plot(LMFitGM, which = 1:2)

# Sqrt Graminoid ANOVA - Analysis used for presentation(s) 
ThesisData$sqrtGM <- sqrt(ThesisData$gm_cover)
hist(ThesisData$sqrtGM)
LMFitsqrtGM <- lm(sqrtGM ~ burn_sev_category * factor(forest_type), data = ThesisData)
anova(LMFitsqrtGM)
par(mfrow = c(1, 2))
plot(LMFitsqrtGM, which = 1:2)

# # log Graminoid ANOVA - not used
# ThesisData$logGM <- log(ThesisData$gm_cover + 1)
# hist(ThesisData$logGM)
# LMFitLogGM <-
#   lm(logGM ~ burn_sev_category * factor(forest_type), data = ThesisData)
# anova(LMFitLogGM)
# par(mfrow = c(1, 2))
# plot(LMFitLogGM, which = 1:2)
```


```{r}
#Tukey pairs Sqrt Graminoids
#Differences by burn severity, but not forest type 
emoutGM <- emmeans(LMFitsqrtGM, ~ burn_sev_category)
emoutGM
pairs(emoutGM)
cld(emoutGM)
```


# Forb ANOVA
```{r, echo=FALSE}
# Forb ANOVA
hist(ThesisData$fb_cover)
LMFitFB <-
  lm(fb_cover ~ burn_sev_category * forest_type, data = ThesisData)
anova(LMFitFB)
par(mfrow = c(1, 2))
plot(LMFitFB, which = 1:2)

# Sqrt Forb ANOVA - Analysis used for presentation(s)
# ThesisData$sqrtFB <- sqrt(ThesisData$fb_cover)
# hist(ThesisData$sqrtFB)
# LMFitsqrtFB <- lm(sqrtFB ~ burn_sev_category * factor(forest_type), data = ThesisData)
# anova(LMFitsqrtFB)
# par(mfrow = c(1, 2))
# plot(LMFitsqrtFB, which = 1:2)

# # log Forb ANOVA - not used
# ThesisData$logFB <- log(ThesisData$fb_cover + 1)
# hist(ThesisData$logFB)
```


```{r}
#Tukey pairs Sqrt Forbs 
#Differences by forest type, but not burn severity 
emoutFB <- emmeans(LMFitFB, ~ factor(forest_type))
emoutFB
pairs(emoutFB)
cld(emoutFB)
```


# Shrub ANOVA
```{r, echo=FALSE}
# # Shrub ANOVA
# hist(ThesisData$sh_cover)
# LMFitSH <-
#   lm(sh_cover ~ burn_sev_category * forest_type, data = ThesisData)
# anova(LMFitSH)
# par(mfrow = c(1, 2))
# plot(LMFitSH, which = 1:2)

# # Sqrt Shrub ANOVA - Analysis used for presentation(s)
# ThesisData$sqrtSH <- sqrt(ThesisData$sh_cover)
# hist(ThesisData$sqrtSH)
# LMFitsqrtSH <- lm(sqrtSH ~ burn_sev_category * factor(forest_type), data = ThesisData)
# anova(LMFitsqrtSH)
# par(mfrow = c(1, 2))
# plot(LMFitsqrtSH, which = 1:2)

# log Shrub ANOVA - not used
ThesisData$logSH <- log(ThesisData$sh_cover + 1)
hist(ThesisData$logSH)
LMFitLogSH <- lm(logSH ~ burn_sev_category * factor(forest_type), data = ThesisData)
anova(LMFitLogSH)
par(mfrow = c(1, 2))
plot(LMFitLogSH, which = 1:2)
```


```{r}
#Tukey pairs Sqrt Shrubs
#Assumptions not met
#Attempt hurble model 
emoutSH <- emmeans(LMFitsqrtSH, ~ burn_sev_category*factor(forest_type))
emoutSH
pairs(emoutSH)
cld(emoutSH)
```


# Tree Seedling ANOVA
```{r, echo=FALSE}
# # Tree ANOVA
# hist(ThesisData$tr_cover)
# LMFitTR <-
#   lm(tr_cover ~ burn_sev_category * forest_type, data = ThesisData)
# anova(LMFitTR)
# par(mfrow = c(1, 2))
# plot(LMFitTR, which = 1:2)
# 
# # sqrt Tree ANOVA - not used
# ThesisData$sqrtTR <- sqrt(ThesisData$tr_cover)
# hist(ThesisData$sqrtTR)
# LMFitsqrtTR <- lm(sqrtTR ~ burn_sev_category * factor(forest_type), data = ThesisData)
# anova(LMFitsqrtTR)

# log Tree ANOVA - Analysis used for presentation(s)
ThesisData$logTR <- log(ThesisData$tr_cover + 1)
hist(ThesisData$logTR)
LMFitLogTR <- lm(logTR ~ burn_sev_category * factor(forest_type), data = ThesisData)
anova(LMFitLogTR)
par(mfrow = c(1, 2))
plot(LMFitLogTR, which = 1:2)
```


```{r}
# Tukey pairs log Tree
#Assumptions not met 
#Attempt hurdble model 
emoutTR <- emmeans(LMFitLogTR, ~ factor(forest_type))
emoutTR
pairs(emoutTR)
cld(emoutTR)
```


#---------------------Seedlings species specific figure-------------------------


# Seedlings long data
```{r}
# Covert seedlings data to long format
SeedlingsLong <- ThesisData %>%
  dplyr::select(forest_type, burn_sev_category, POTR:RONE, PIPO:PIEN) %>% 
  tidyr::pivot_longer(
    cols=c(POTR:RONE,PIPO:PIEN),names_to = "Species", values_to = "Density") %>%
  mutate(forest_type = factor(
    forest_type,
    levels = c("Lower Montane", "Upper Montane","Subalpine"))) %>%
      mutate(Species = factor(Species,
                                    levels = c("PIPO","PSME","ABCO","PICO","PIFL", "PIAR","ABLA","PIEN", "QUGA", "RONE","POTR")))
```


# Facet axis removing labels for species not present in forest types 
```{r}
#Omit na values
SeedlingsLong2 <- SeedlingsLong %>% 
  na.omit
```


# Seedlings plot 
```{r}
# Facet wrap by forest type - seedlings plot
ggplot(SeedlingsLong2,
       aes(x = Species, y = Density, fill = burn_sev_category)) +
  geom_boxplot() + ylim(0, 9500) +
  facet_wrap(~ forest_type, scales="free_x") +
  theme_classic() +
  labs(x = "Species",
       y = "Seedlings per hectare",
       fill = "Burn Severity",
       title = "Seedling Density") +
  theme(legend.position = "bottom",
        legend.key.width = unit(1, "cm")) +
  theme_few() +
  scale_fill_manual(
    name = "Burn Severity",
    labels = c("Low", "High"), values=c("#fd8d3c","#bd0026")) +
  theme(
    axis.text.x = element_text(
      color = "black",
      size = 6,
      angle = 45,
      vjust = .8,
      hjust = .8
    ))
```


#----------- Conifer/deciduous figure-----------------------------------------------


# Conifer/decid transformations and distributions
```{r, echo=FALSE}
# Conifer/decid transformations and distributions
ThesisData$sqrtConifer <- sqrt(ThesisData$conifer_total)
ThesisData$sqrtDeciduous <- sqrt(ThesisData$deciduous_total)
ThesisData$logConifer <- log(ThesisData$conifer_total + 1)
ThesisData$logDeciduous <- log(ThesisData$deciduous_total + 1)
par(mfrow = c(3, 3))
hist(ThesisData$conifer_total)
hist(ThesisData$deciduous_total)
hist(ThesisData$sqrtConifer)
hist(ThesisData$sqrtDeciduous)
hist(ThesisData$logConifer)
hist(ThesisData$logDeciduous)
```


# Conifer/decid long data 
```{r}
# Covert conifer/decid data to long format
DecidConiferLong1 <- ThesisData %>%
  dplyr::select(forest_type,
         burn_sev_category,
         conifer_total,
         deciduous_total) %>%
  tidyr::pivot_longer(
    cols = c("conifer_total", "deciduous_total"),
    names_to = "TreeGroup",
    values_to = "Density") %>%
  mutate(forest_type = factor(
    forest_type,
    levels = c("Lower Montane", "Upper Montane","Subalpine"))) %>%
  mutate(burn_sev_category = factor(burn_sev_category, levels = c("Low","High")))%>%
  mutate(TreeGroup = factor(TreeGroup, labels = c("Conifer","Deciduous")))
```


# Conifer/Decid plot 
```{r}
# Facet wrap by forest type - conifer/decid plot
ggplot(DecidConiferLong1,
       aes(x = TreeGroup, y = Density, fill = burn_sev_category)) +
  geom_boxplot() + ylim(0, 28500) +
  facet_wrap(~ forest_type) +
  theme_classic() +
  labs(x = "",
       y = "Seedlings per hectare",
       fill = "Burn Severity",
       title = "Conifer versus Deciduous Regeneration") +
  theme(legend.position = "bottom",
        legend.key.width = unit(1, "cm")) +
  theme_few() +
  scale_fill_manual(
    name = "Burn Severity",
    labels = c("Low", "High"),
    values = c("#fd8d3c","#bd0026")
  ) +
  theme(
    axis.text.x = element_text(
      color = "black",
      size = 8,
      angle = 0,
      vjust = .8,
      hjust = .8
    ) 
  ) 
```


#------------Conifer/decid analyses---------------------------------------------


## Deciduous density 


# Two way ANOVA deciduous density only, sqrt transform
```{r}
# Two way ANOVA deciduous, sqrt transform
LMFitDecidsqrt <- lm(sqrtDeciduous ~ burn_sev_category * forest_type, data = ThesisData)
anova(LMFitDecidsqrt)
par(mfrow = c(1, 2))
plot(LMFitDecidsqrt, which = 1:2)
```


```{r}
# LMFitDecid <- lm(log(deciduous_total+.00001) ~burn_sev_category* forest_type, data = ThesisData)
# anova(LMFitDecid)
# par(mfrow = c(1, 2))
# plot(LMFitDecid, which = 1:2)
```


# Tukey pairs Deciduous Density - Forest Type
```{r}
#Tukey pairs Deciduous density 
emoutDecid <- emmeans(LMFitDecidsqrt, ~forest_type)
pairs(emoutDecid)
cld(emoutDecid)
```


## Conifer density 


# Zero-inflated Poisson Conifer 
```{r}
ThesisData$conifer_total<-as.integer(ThesisData$conifer_total)

ZpoiConiferFit <- zeroinfl(conifer_total ~ burn_sev_category * forest_type, data=ThesisData)
summary(ZpoiConiferFit)
```



# Zero-inflated Poisson Deciduous 
```{r}
ThesisData$deciduous_total<-as.integer(ThesisData$deciduous_total)

ZpoiDecidFit <- zeroinfl(deciduous_total ~ burn_sev_category * forest_type, data=ThesisData)
summary(ZpoiDecidFit)
```



















